Pulmonary Exacerbations in CF

### 2024-11-07 Scott Sagel/PEx in CF/analysis_for_yaron.qmd
added white vs. non-white

```{r setup}
#| include: false
library(dplyr)
library(tidyr)
library(lubridate)
library(gtsummary)
library(ggpubr)
library(rspiro)
library(patchwork)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Scott Sagel/PEx in CF",
  "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Scott Sagel/PEx in CF",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Scott Sagel/PEx in CF"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/GitHub",
  "Windows" = "C:/Users/timvigers/GitHub",
  "Linux" = "/home/timvigers/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
# Colors for plots
yellow <- "#f8b80a"
red <- "#e30022"
purple <- "#7b2670"
blue <- "#022451"
```

@@

@@

df$no_bact <- apply(df, 1, function(r) {
  r <- as.numeric(r[paste0("admit_cx_results___", 1:8)])
  all(r == 0)
})
df$no_bact <- factor(df$no_bact, levels = c(F, T), labels = c("No", "Yes"))
# Change in FEV1 from admission to discharge
df$ppfev1_change <- df$ppfev1_discharge - df$ppfev1_admit
# Format fungi, NTM, etc. columns
df <- df %>%
  mutate(
    across(
      c(
        fungi___1:fungi___5, ntm___1:ntm___3, viruses___1:viruses___5,
        admit_cx_results___1:admit_cx_results___9
      ),
      ~ factor(as.numeric(.x),
        levels = 0:1,
        labels = c("Not detected", "Detected")
      )
    )
  )
# Labels
labels <- list(
  "Age group at admission" = "age_group",
  "Age at admission" = "age_at_admit", "Sex" = "sex",
  "Race" = "race", "Ethnicity" = "ethnicity",
  "Azithromycin" = "azithro", "Inhaled antibiotics" = "inhaled_abx",
  "Systemic steroids" = "steroids_system",
  "Source of respiratory culture at admit" = "admit_cx_source",
  "Aspergillus fumigatus" = "fungi___1",
  "Aspergillus (other species)" = "fungi___2",
  "Scedosporium apiospermum" = "fungi___3",
  "Candidia (any species)" = "fungi___4",
  "Other fungus" = "fungi___5",
  "Any fungi?" = "any_fungi",
  "NTM testing performed?" = "ntm_test",
  "Fungal testing performed?" = "fungal_test",
  "Viral testing performed?" = "viral_test",
  "Viral testing results" = "viral_test_results",
  "Influenza A or B" = "viruses___1",
  "Respiratory Syncytial Virus" = "viruses___2",
  "Rhinovirus/enterovirus" = "viruses___3",
  "SARS CoV-2" = "viruses___4",
  "Other virus" = "viruses___5",
  "Any virus?" = "any_virus",
  "M. abscessus species" = "ntm___1",
  "M. avium complex species" = "ntm___2",
  "Other NTM species" = "ntm___3",
  "Any NTM?" = "any_ntm",
  "A. xylosoxidans" = "admit_cx_results___1",
  "B. cepacia complex" = "admit_cx_results___2",
  "H. influenzae" = "admit_cx_results___3",
  "MRSA" = "admit_cx_results___4",
  "Methicillin-susceptible Staph aureus" = "admit_cx_results___5",
  "P. aeruginosa" = "admit_cx_results___6",
  "P. aeruginosa mucoid" = "admit_cx_results___7",
  "S. maltophilia" = "admit_cx_results___8",
  "Other bacteria" = "admit_cx_results___9",
  "Any bacteria?" = "any_bact",
  "No CF bacteria" = "no_bact",
  "On CFTR Modulator?" = "cftr_mod",
  "CFTR Modulator Type" = "cftr_mod_spec",
  "Site" = "redcap_data_access_group",
  "Length of Admission" = "days_admit",
  "Systemic Steroids" = "steroids_system",
  "Oral Steroids" = "steroids_oral",
  "IV Steroids" = "steroids_iv",
  "Baseline FEV1pp" = "ppfev1_baseline",
  "Admission FEV1pp" = "ppfev1_admit",
  "Discharge FEV1pp" = "ppfev1_discharge",
  "Change in mean FEV1pp from admit to discharge" = "ppfev1_change"
)
labels <- as.list(setNames(names(labels), labels))
# Format
df <- df %>%
  select(
    sid,
    redcap_data_access_group, year, sex, age_group, age_at_admit, race,
    ethnicity, Genotype, Insurance, cftr_mod, cftr_mod_spec,
    Trikafta, azithro, inhaled_abx, steroids_system, admit_cx_source,
    ppfev1_baseline, ppfev1_admit, ppfev1_discharge, ppfev1_change,
    days_admit, fungal_test, fungi___1:fungi___5, any_fungi,
    ntm_test, ntm___1:ntm___3, any_ntm,
    viral_test, viruses___1:viruses___5, any_virus,
    admit_cx_results___1:admit_cx_results___9, any_bact, no_bact,
    contains("steroids")
  ) %>%
  filter(!is.na(year))
```

@@

```{r}
#| label: tbl-table-1
#| tbl-cap: Participant characteristics overall
df %>%
  group_by(sid) %>%
  filter(row_number() == 1) %>%
  ungroup() %>%
  select(redcap_data_access_group, sex, race, ethnicity, Genotype) %>%
  tbl_summary(missing_text = "(Missing)", label = labels)
```

```{r}
#| label: tbl-table-1-year
#| tbl-cap: Participant characteristics by year
df %>%
  group_by(sid, year) %>%
  filter(row_number() == 1) %>%
  ungroup() %>%
  select(year, redcap_data_access_group, sex, race, ethnicity, Genotype) %>%
  tbl_summary(by = year, missing_text = "(Missing)", label = labels)
```

```{r}
#| label: tbl-table-1-indep
#| tbl-cap: Participant characteristics at admission by year
df %>%
  select(redcap_data_access_group:days_admit) %>%
  tbl_summary(
    by = year, missing_text = "(Missing)", label = labels,
    statistic =
      list(
        all_continuous() ~ "{median} [{p25}, {p75}]",
        c(
          age_at_admit, ppfev1_baseline, ppfev1_admit, ppfev1_discharge,
          ppfev1_change
        ) ~
          "{mean} ({sd})"
      )
  ) %>%
  add_p(
    test =
      list(
        c(age_at_admit, ppfev1_baseline, ppfev1_admit, ppfev1_discharge) ~
          "t.test"
      ),
    test.args = all_tests("fisher.test") ~ list(simulate.p.value = T)
  ) %>%
  separate_p_footnotes()
```

>>>>

```{r setup}
#| include: false
library(dplyr)
library(tidyr)
library(lubridate)
library(gtsummary)
library(ggpubr)
library(rspiro)
library(patchwork)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Scott Sagel/PEx in CF",
  "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Scott Sagel/PEx in CF",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Scott Sagel/PEx in CF"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/GitHub",
  "Windows" = "C:/Users/timvigers/Documents/GitHub",
  "Linux" = "/home/timvigers/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
# Colors for plots
yellow <- "#f8b80a"
red <- "#e30022"
purple <- "#7b2670"
blue <- "#022451"
```

@@

df$race_wnw <- factor(df$race,
    "White", "Black or African-American", "Other", "Asian",
    "Unknown/not reported"
  labels = c("White", "Non-White", "Non-White", "Non-White", "Non-White")

@@

df$no_bact <- apply(df, 1, function(r) {
  r <- as.numeric(r[paste0("admit_cx_results___", 1:8)])
  all(r == 0)
})
df$no_bact <- factor(df$no_bact, levels = c(F, T), labels = c("No", "Yes"))
# Change in FEV1 from admission to discharge
df$ppfev1_change <- df$ppfev1_discharge - df$ppfev1_admit
# Format fungi, NTM, etc. columns
df <- df %>%
  mutate(
    across(
      c(
        fungi___1:fungi___5, ntm___1:ntm___3, viruses___1:viruses___5,
        admit_cx_results___1:admit_cx_results___9
      ),
      ~ factor(as.numeric(.x),
        levels = 0:1,
        labels = c("Not detected", "Detected")
      )
    )
  )
# Labels
labels <- list(
  "Age group at admission" = "age_group",
  "Age at admission" = "age_at_admit", "Sex" = "sex",
  "Race" = "race", "Race (White vs. Non-White)" = "race_wnw",
  "Ethnicity" = "ethnicity", "Azithromycin" = "azithro",
  "Inhaled antibiotics" = "inhaled_abx",
  "Systemic steroids" = "steroids_system",
  "Source of respiratory culture at admit" = "admit_cx_source",
  "Aspergillus fumigatus" = "fungi___1",
  "Aspergillus (other species)" = "fungi___2",
  "Scedosporium apiospermum" = "fungi___3",
  "Candidia (any species)" = "fungi___4",
  "Other fungus" = "fungi___5",
  "Any fungi?" = "any_fungi",
  "NTM testing performed?" = "ntm_test",
  "Fungal testing performed?" = "fungal_test",
  "Viral testing performed?" = "viral_test",
  "Viral testing results" = "viral_test_results",
  "Influenza A or B" = "viruses___1",
  "Respiratory Syncytial Virus" = "viruses___2",
  "Rhinovirus/enterovirus" = "viruses___3",
  "SARS CoV-2" = "viruses___4",
  "Other virus" = "viruses___5",
  "Any virus?" = "any_virus",
  "M. abscessus species" = "ntm___1",
  "M. avium complex species" = "ntm___2",
  "Other NTM species" = "ntm___3",
  "Any NTM?" = "any_ntm",
  "A. xylosoxidans" = "admit_cx_results___1",
  "B. cepacia complex" = "admit_cx_results___2",
  "H. influenzae" = "admit_cx_results___3",
  "MRSA" = "admit_cx_results___4",
  "Methicillin-susceptible Staph aureus" = "admit_cx_results___5",
  "P. aeruginosa" = "admit_cx_results___6",
  "P. aeruginosa mucoid" = "admit_cx_results___7",
  "S. maltophilia" = "admit_cx_results___8",
  "Other bacteria" = "admit_cx_results___9",
  "Any bacteria?" = "any_bact",
  "No CF bacteria" = "no_bact",
  "On CFTR Modulator?" = "cftr_mod",
  "CFTR Modulator Type" = "cftr_mod_spec",
  "Site" = "redcap_data_access_group",
  "Length of Admission" = "days_admit",
  "Systemic Steroids" = "steroids_system",
  "Oral Steroids" = "steroids_oral",
  "IV Steroids" = "steroids_iv",
  "Baseline FEV1pp" = "ppfev1_baseline",
  "Admission FEV1pp" = "ppfev1_admit",
  "Discharge FEV1pp" = "ppfev1_discharge",
  "Change in mean FEV1pp from admit to discharge" = "ppfev1_change"
)
labels <- as.list(setNames(names(labels), labels))
# Format
df <- df %>%
  select(
    sid,
    redcap_data_access_group, year, sex, age_group, age_at_admit, race,
    race_wnw, ethnicity, Genotype, Insurance, cftr_mod, cftr_mod_spec,
    Trikafta, azithro, inhaled_abx, steroids_system, admit_cx_source,
    ppfev1_baseline, ppfev1_admit, ppfev1_discharge, ppfev1_change,
    days_admit, fungal_test, fungi___1:fungi___5, any_fungi,
    ntm_test, ntm___1:ntm___3, any_ntm,
    viral_test, viruses___1:viruses___5, any_virus,
    admit_cx_results___1:admit_cx_results___9, any_bact, no_bact,
    contains("steroids")
  ) %>%
  filter(!is.na(year))
```

@@

```{r}
#| label: tbl-table-1
#| tbl-cap: Participant characteristics overall
df %>%
  group_by(sid) %>%
  filter(row_number() == 1) %>%
  ungroup() %>%
  select(redcap_data_access_group, sex, race, race_wnw, ethnicity, Genotype) %>%
  tbl_summary(missing_text = "(Missing)", label = labels)
```

```{r}
#| label: tbl-table-1-year
#| tbl-cap: Participant characteristics by year
df %>%
  group_by(sid, year) %>%
  filter(row_number() == 1) %>%
  ungroup() %>%
  select(
    year, redcap_data_access_group, sex, race, race_wnw, ethnicity, Genotype
  ) %>%
  tbl_summary(by = year, missing_text = "(Missing)", label = labels) %>%
  add_p() %>%
  separate_p_footnotes()
```

```{r}
#| label: tbl-table-1-indep
#| tbl-cap: Participant characteristics at admission by year
df %>%
  select(redcap_data_access_group:days_admit) %>%
  tbl_summary(
    by = year, missing_text = "(Missing)", label = labels,
    statistic =
      list(
        all_continuous() ~ "{median} [{p25}, {p75}]",
        c(
          age_at_admit, ppfev1_baseline, ppfev1_admit, ppfev1_discharge
        ) ~
          "{mean} ({sd})"
      )
  ) %>%
  add_p(
    test =
      list(
        c(age_at_admit, ppfev1_baseline, ppfev1_admit, ppfev1_discharge) ~
          "t.test"
      ),
    test.args = all_tests("fisher.test") ~ list(simulate.p.value = T)
  ) %>%
  separate_p_footnotes()
```
